Untitled

### 2020-11-20 ms/manuscript.Rmd
Add vapor pressure decifit (vpd) to PGLS analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
library(jntools)
```

```{r load-data}
loadd(
  list = c(
    "climate",
    "filmy_species",
    "moorea_sites",
    "phylosig",
    "range_types",
    "recovery_species_means",
    "t_test_results"
  )
)
```

@@

```{r dt-stats}
# Count number of species included in DT test by generation
n_dt_species_gameto <- recovery_species_means %>% 
  count(species, generation) %>% 
  count(generation) %>% 
  filter(generation == "gameto") %>% 
  pull(n) %>% english2

n_dt_species_gameto <- recovery_species_means %>% 
  count(species, generation) %>% 
  count(generation) %>% 
  filter(generation == "sporo") %>% 
  pull(n) %>% english2

n_dt_species_both <- recovery_species_means %>% 
  count(species, generation) %>% 
  group_by(species) %>% 
  summarize(n_generation = n_distinct(generation), .groups = "drop") %>%
  filter(n_generation == 2) %>%
  pull(species) %>%
  n_distinct() %>%
  english2

@@

Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_gameto` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on these species after confirming their lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

@@

@@

Light responses were measured in sporophytes of 18 species and gametophytes of 11 species, including nine species with both life phases (Appendix S4).
Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig. 5).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
142 ± 46 µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at 51 ± 22 µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; 8.1 ± 2.7 µmol·m^--2^ ·s^--1^) than gametophytes (3.9 ± 0.9 µmol·m^--2^ ·s^--1^).
The GLMMs showed a significant effect of generation, but neither range type nor growth habit, on PPFD~95%~ and ETR~max~ (Table 3).

### Environmental water potential and DT

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide", cache = FALSE)
```

```{r load-data}
loadd(
  list = c(
    "climate",
    "filmy_species",
    "moorea_sites",
    "phylosig",
    "range_types",
    "combined_species_means",
    "t_test_results"
  )
)
```

@@

```{r dt-stats}
# Count number of species included in DT test by generation
n_dt_species_sporo <- combined_species_means %>% 
  filter(!is.na(recover_mean)) %>% 
  count(species, generation) %>% 
  count(generation) %>% 
  filter(generation == "sporo") %>% 
  pull(n) %>% 
  english2

n_dt_species_gameto <- combined_species_means %>% 
  filter(!is.na(recover_mean)) %>%
  count(species, generation) %>% 
  count(generation) %>% 
  filter(generation == "gameto") %>% 
  pull(n) %>% 
  english2

n_dt_species_both <- combined_species_means %>% 
  filter(!is.na(recover_mean)) %>% 
  count(species, generation) %>% 
  group_by(species) %>% 
  summarize(n_generation = n_distinct(generation), .groups = "drop") %>%
  filter(n_generation == 2) %>%
  pull(species) %>%
  n_distinct() %>%
  english2

@@

Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases (Appendix S3).
A general correlation of DT with habitat was observed in sporophytes (Fig. 4).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 MPa (more stringent tests were not conducted on these species after confirming their lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

@@

```{r photosyn-stats}
# Count number of species included in light data by generation
n_light_species_sporo <- combined_species_means %>% 
  filter(!is.na(etr_mean), !is.na(par_mean)) %>%
  count(species, generation) %>% 
  count(generation) %>% 
  filter(generation == "sporo") %>% 
  pull(n) %>% 
  english2

n_light_species_gameto <- combined_species_means %>% 
  filter(!is.na(etr_mean), !is.na(par_mean)) %>%
  count(species, generation) %>% 
  count(generation) %>% 
  filter(generation == "gameto") %>% 
  pull(n) %>% 
  english2

n_light_species_both <- combined_species_means %>% 
  filter(!is.na(etr_mean), !is.na(par_mean)) %>% 
  count(species, generation) %>% 
  group_by(species) %>% 
  summarize(n_generation = n_distinct(generation), .groups = "drop") %>%
  filter(n_generation == 2) %>%
  pull(species) %>%
  n_distinct() %>%
  english2

# Calculate grand means by generation for light responses
# (so mean is the mean of species means, and sd is sd of species means)
light_overall_means <-
combined_species_means %>%
  group_by(generation) %>%
  summarize(
    etr = mean(etr_mean, na.rm = TRUE),
    etr_sd = sd(etr_mean, na.rm = TRUE),
    par = mean(par_mean, na.rm = TRUE),
    par_sd = sd(par_mean, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(across(where(is.numeric), ~round_t(.x, 1))) %>%
  mutate(
    etr_pretty = glue("{etr} ± {etr_sd}"),
    par_pretty = glue("{par} ± {par_sd}"),
  )

# Extract values by response and generation for citing in MS
mean_par_sporo <- light_overall_means %>% filter(generation == "sporo") %>% pull(par_pretty)
mean_par_gameto <- light_overall_means %>% filter(generation == "gameto") %>% pull(par_pretty)
mean_etr_sporo <- light_overall_means %>% filter(generation == "sporo") %>% pull(etr_pretty)
mean_etr_gameto <- light_overall_means %>% filter(generation == "gameto") %>% pull(etr_pretty)
```

Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases (Appendix S4).
Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig. 5).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).
The GLMMs showed a significant effect of generation, but neither range type nor growth habit, on PPFD~95%~ and ETR~max~ (Table 3).

### Relationships of water potential and range size with DT

No significant relationship was detected between range size of gametophytes and DT.

@@
